Reaching is Better When You Get What You Want: Realtime Feedback of Intended Reaching Trajectory Despite an Unstable Environment

### 2015-06-08 Paper1rev3/paper1rev3.tex
Updated for submitted version

@@

The simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting. More accurate approaches would provide us with many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result. Our results point to the possibility of the impedance changing during motion, perhaps as a function of time.

@@

%\section*{Supporting Information Legends}
%
% Please enter your Supporting Information captions below in the following format:
%\item{\bf Figure SX. Enter mandatory title here.} Enter optional descriptive information here.
% 
%\begin{description}
%\item {\bf}
%\item {\bf}
%\end{description}

>>>>

Since trajectories were not controlled experimentally, our tests rely heavily on the assumption that disturbed intentions initially mirror undisturbed intentions. Therefore, it is not completely certain that subjects intended to move along a straight line to the target. However, this study examined straight line motions because it is well understood that people tend to repetitively attempt to reach in a straight line towards a target \cite{hodgson2000model}. Also, even when subjects are persistently perturbed by a force field, they recover their motions to a straight line with repetitive experience \cite{shadmehr1994adaptive}. Nevertheless, it may be better to employ tasks with a more specific trajectory in order to validate the intended path. Unfortunately, even such an intended trajectory may not be strictly invariant during external disturbance because some displacements may cause intent to be recalculated by the nervous system.

@@

The simple linear scaling of Burdet et al.'s stiffness model \cite{burdet2006stability} facilitated hypothesis testing, but was inherently limiting. More accurate approaches would provide us with many free parameters through which to determine the outcome of the eventual extraction and obtain a spurious favorable result. Like others who have examined large force disturbances, we find lower stiffnesses (our $c$ parameter) than might be expected \cite{crevecoeur2014beyond} (Table in S1 Table). This may be due to differences in tasks \cite{gomi1998task} and/or the dependency of muscle impedance on perturbation size \cite{van2010vivo}. Our results also point to the possibility of the impedance changing during motion, perhaps as a function of time.

@@

\section*{Supporting Information Legends}

\begin{description}
\item{\bf Table S1. Subject-specific parameters.} Gross body mass was self-reported. Upper arm length, forearm length, and shoulder position in the robot's coordinate system were measured \textit{in situ}. Feedback torque gain was fit as described in our methods section using the intermittently presented white noise force disturbance. This gain is relative to the total feedback torque output (stiffness, damping, and reflexes) of the model of Burdet et al.\cite{burdet2006stability}.
\end{description}

### 2015-06-08 Paper2frontiers/frontiers.tex
Updated for submitted version

\subsection{Human Subjects}
The human data trajectories analyzed here are drawn from sixteen subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki. fourteen male and two female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated. Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.

\subsection{Experimental Design}
We chose 15 centimeters long, center-out reaches of the right arm to visually-presented targets chosen at $60^\circ$ intervals about that central point. Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each of five blocks of 96 reaches each. During blocks 2 through 4, subjects experienced filtered white noise forces drawn from a white noise generator at 1000 Hertz with flat power spectral density of 1 Newton, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ radians per second. During block 3, cursor position indicated deduced intent. During all other blocks, the cursor position indicated hand position (Figure 2A).

>>>>

\subsection{Human Subjects}
The human data trajectories analyzed here are drawn from sixteen subjects who gave informed consent in accordance with Northwestern University Institutional Review Board, which specifically approved this study and follows the principles expressed in the Declaration of Helsinki. Fourteen male and two female right-handed subjects (ages 21 to 30) performed the reaches with their right arm and were not compensated. Subjects' arm segment lengths were directly measured \textit{in situ} while body mass was self-reported.

\subsection{Experimental Design}
We chose center-out reaches of the right arm to visually-present targets 15 centimeters from the center chosen at $60^\circ$ intervals. Target selection was carried out pseudorandomly such that each outer target was visited 16 times in each of five blocks of 96 reaches each. During blocks 2 through 4, subjects experienced filtered white noise forces drawn from a white noise generator at 1000 Hertz with flat power spectral density of 1 Newton, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ radians per second. During block 3, cursor position indicated deduced intent. During all other blocks, the cursor position indicated hand position (Figure 2A).
